---
title: "R Notebook"
output: html_notebook
---

This document describes drawing transect lines within SDB for habitat sampling.

First load necessary libraries.
```{r}
rm(list=ls())
ifelse(Sys.info()[1] == 'Linux',
       source('~/Documents/R/tools/TomosFunctions.R'),
       source('~/R/tools/TomosFunctions.R'))

library(rgdal)
library(sp)
library(geosphere)
library(rgeos)
library(tidyverse)

dlines <- 25    # distance between lines in m

```

The following code was modified from a script, which was used to draw aerial transect lines for the Southern California Bight (CaptureArea.R in SCB_AerialSurvey project).

First, bring in a shapefile for the bay and define a dataframe. 
```{r}

# Bring in a map of SDB:
SDBay.gis <- spTransform(readOGR(dsn = "GISfiles",
                                            layer = "sd_bay",
                                            verbose = FALSE),
                             CRS("+proj=longlat +datum=WGS84"))

SDBay.df <- broom::tidy(SDBay.gis)
```


First line is defined by selecting the northern most latitude and finding the longitude of the shore lines (both E and W). To do that, we use... over in sp package?
```{r}


# find N most E end along the shore line:

N.most <- data.frame(long = -117.1480, lat = 32.6863)
S.most <- data.frame(long = -117.1115, lat = 32.6059)

rhumb.bear <- bearingRhumb(S.most, N.most)

pt2.df <- data.frame(X = N.most$long, Y = N.most$lat)

## draw transect lines:
# find the distance between the N and S tips of the study area in m:
dist_NS <- distGeo(N.most, S.most)
nlines <- ceiling(dist_NS/dlines)
k<-2
for (k in 2:nlines){
  # find two points on the next line
  # Offshore
  pt2.df[k,] <- destPointRhumb(pt2.df[k-1,], rhumb_bear-90, dlines*1000)
  # Inshore - go long enough because it will be truncated later.
  pt1.df[k,] <- destPointRhumb(pt2.df[k,], rhumb_bear-180, nm2km(60)*1000)
  #pt1.df[k,] <- destPointRhumb(pt1.df[k-1,], rhumb_bear-90, dlines*1000)

}

pt1.df.sp <- latlon2sp(in.df = pt1.df, center.UTM = center.UTM)
pt2.df.sp <- latlon2sp(in.df = pt2.df, center.UTM = center.UTM)

# find intersections between lines and the outside boundary line:
# Create Lines objects and put them in a list
lines.list <- vector(mode = 'list', length = (nlines-1))
k <- 2
for (k in 2:nlines){
  line1 <- as.matrix(rbind(pt1.df.sp@coords[k,],
                           pt2.df.sp@coords[k,]))
  lines.list[[k-1]] <- Lines(list(Line(line1)),
                             ID = paste0('line', k))
}

# Create SpatialLines object
line1.Sp <- SpatialLines(lines.list,
                         proj4string = CRS("+proj=utm +zone=10 ellps=WGS84"))

# use gIntersection to find intersections between lines and the polygon
# This was a headache... found some code, which is saved in gIntersection_example.R
edges <- coordinates(gIntersection(line1.Sp, search.area.Sp.poly))

# create two empty data frames for begin and end points
transect.begin.lines.Sp <- data.frame(X = NA,
                                      Y = NA)
transect.end.lines.Sp <- data.frame(X = NA,
                                    Y = NA)

# extract the edge points into the dataframes
for (k in 1:length(edges[[1]])){
  transect.begin.lines.Sp[k, 'X'] <- edges[[1]][[k]][1, 'x']
  transect.begin.lines.Sp[k, 'Y'] <- edges[[1]][[k]][1, 'y']
  transect.end.lines.Sp[k, 'X'] <- edges[[1]][[k]][2, 'x']
  transect.end.lines.Sp[k, 'Y'] <- edges[[1]][[k]][2, 'y']

}

# flip the data frames upside down so they are numbered from north to south
# apply converts dataframes into matrices so convert them bacck to
# dataframes. Add the first line - the northern boundary line first.
transect.begin.lines.Sp <- as.data.frame(apply(rbind(transect.begin.lines.Sp,
                                                     pt1.df.sp@coords[1, ]), 2, rev))
transect.end.lines.Sp <- as.data.frame(apply(rbind(transect.end.lines.Sp,
                                                   pt2.df.sp@coords[1, ]), 2, rev))

# convert these into spatial objects with appropriate projections
coordinates(transect.begin.lines.Sp) <- c('X', 'Y')
proj4string(transect.begin.lines.Sp) <- CRS("+proj=utm +zone=10 ellps=WGS84")
coordinates(transect.end.lines.Sp) <- c('X', 'Y')
proj4string(transect.end.lines.Sp) <- CRS("+proj=utm +zone=10 ellps=WGS84")

# convert them into lat/lon
transect.begin.lines.latlon <- spTransform(transect.begin.lines.Sp,
                                           CRS("+proj=longlat +datum=WGS84"))

transect.end.lines.latlon <- spTransform(transect.end.lines.Sp,
                                           CRS("+proj=longlat +datum=WGS84"))

# combine them in one data frame along with track lines:
transect.lines.latlon <- data.frame(lineID = 1:nrow(transect.begin.lines.Sp@coords),
                                    latInshore = transect.begin.lines.latlon@coords[,'Y'],
                                    lonInshore = transect.begin.lines.latlon@coords[,'X'],
                                    latOffshore = transect.end.lines.latlon@coords[, 'Y'],
                                    lonOffshore = transect.end.lines.latlon@coords[, 'X'])

# provide decimal degrees and decimal minutes in the output:
transect.lines.latlon$degLatInshore <- trunc(transect.lines.latlon$latInshore)
transect.lines.latlon$minLatInshore <- abs((transect.lines.latlon$latInshore - trunc(transect.lines.latlon$latInshore)) * 60)
transect.lines.latlon$degLonInshore <- trunc(transect.lines.latlon$lonInshore)
transect.lines.latlon$minLonInshore <- abs((transect.lines.latlon$lonInshore - trunc(transect.lines.latlon$lonInshore)) * 60)

transect.lines.latlon$degLatOffshore <- trunc(transect.lines.latlon$latOffshore)
transect.lines.latlon$minLatOffshore <- abs((transect.lines.latlon$latOffshore - trunc(transect.lines.latlon$latOffshore)) * 60)
transect.lines.latlon$degLonOffshore <- trunc(transect.lines.latlon$lonOffshore)
transect.lines.latlon$minLonOffshore <- abs((transect.lines.latlon$lonOffshore - trunc(transect.lines.latlon$lonOffshore)) * 60)

transect.lines.latlon$distInNM <- km2nm(distGeo(transect.begin.lines.latlon@coords,
                                                transect.end.lines.latlon@coords)/1000)

transect.lines.latlon$dist2MYF <- km2nm(distGeo(transect.begin.lines.latlon@coords,
                                                MYF)/1000)
transect.lines.latlon$dist2CRQ <- km2nm(distGeo(transect.begin.lines.latlon@coords,
                                                CRQ)/1000)
transect.lines.latlon$dist2RNM <- km2nm(distGeo(transect.begin.lines.latlon@coords,
                                                RNM)/1000)

# the last one is in the bay - remove: (this may not happen when dline is changed)
#transect.lines.latlon <- transect.lines.latlon[-12,]

# make a plot
p1 <- ggplot() +
  geom_polygon(fill = land.color,
               data = all.coast.line.df,
               aes(x = X, y = Y, group = idx))  +
  geom_polygon(data = search.area,
               aes(x = X, y = Y),
               fill = NA, size = 1.2,
               alpha = 0.6,
               color='black') +
  geom_segment(data = transect.lines.latlon,
               aes(x = lonInshore, y = latInshore,
                   xend = lonOffshore, yend = latOffshore),
               size = 0.9) +
  geom_text(data = transect.lines.latlon,
            aes(x = lonOffshore,
                y = latOffshore, label = lineID),
            color = 'red')+
  coord_map() +
  xlab('') + ylab('')

if (save.data){
  saveRDS(search.area,
          file = paste0("RDSfiles/SearchArea_", Sys.Date(), ".rds"))
  saveRDS(transect.lines.latlon,
          file = paste0("RDSfiles/TransectLines_", Sys.Date(), ".rds"))
  saveRDS(all.coast.line.df,
          file = paste0("RDSfiles/CoastLine_", Sys.Date(), ".rds"))
}


if (save.image) {
  ggsave(plot = p1,
         dpi = dpi,
         file = paste0('Figures/tracklines_SC_', SanClemente, '_',
                       dlines, 'km_', Sys.Date(), '.png'))

  write.table(signif(transect.lines.latlon, digits = 6), sep = ',',
            file = paste0('Data/tracklines_SC_', SanClemente, '_',
                          dlines, 'km_', Sys.Date(), '.csv'),
            append = F, quote = F, row.names = F, col.names = T)
}

```

