---
title: "SDB E-W transect lines"
output: html_notebook
---

This document describes drawing transect lines within SDB for UAS sampling.

First load necessary libraries.
```{r}
rm(list=ls())
save.image <- F
# ifelse(Sys.info()[1] == 'Linux',
#        source('~/Documents/R/tools/TomosFunctions.R'),
#        source('~/R/tools/TomosFunctions.R'))

library(rgdal)
library(sp)
library(geosphere)
library(rgeos)
library(tidyverse)

dlines <- 300    # distance between lines in m
min.length <- 500  # minimum line length in m

N.most.lat <- 32.622 #32.650
S.most.lat <- 32.604

E.end <- -116
W.end <- -118


water.color <- "lightblue"
background.color <- "darkgray"
eelgrass.color <- "lightgreen"

```

The following code was modified from a script, which was used to draw aerial transect lines for the Southern California Bight (CaptureArea.R in SCB_AerialSurvey project).

First, bring in a shapefile for the bay and define a dataframe. 
```{r}

# Bring in a map of SDB:
SDBay.gis <- spTransform(readOGR(dsn = "GISfiles",
                                 layer = "sd_bay",
                                 verbose = FALSE),
                         CRS("+proj=longlat +datum=WGS84"))

SDBay.df <- broom::tidy(SDBay.gis)
```


First line is defined by selecting the northern most latitude for transect lines and finding the longitude of the shore lines (both E and W). This is completed by using gIntersection.  The while loop continues until the new latitude is south of the southern limit. 
```{r}
# N/S most latitudes for transect lines
# create a line along this latitude (longitude values are arbitrary: wide enough to cover SDB):
N.most.line <- SpatialLines(list(Lines(list(Line(data.frame(long = c(W.end, E.end),
                                                            lat = c(N.most.lat, N.most.lat)))), 
                                       ID = "Reference")),
                            proj4string = CRS("+proj=longlat +datum=WGS84"))

# ends of the northern most line
N.coords <- coordinates(gIntersection(SDBay.gis, N.most.line))
N.coords.df <- data.frame(N.coords)
#NE.coords <- N.coords.df[2,]

E.coords <- data.frame(long = N.coords.df[2, "x"],
                       lat = N.coords.df[2, "y"])

W.coords <- data.frame(long = N.coords.df[1, "x"],
                       lat = N.coords.df[1, "y"])

lat <- E.coords$lat

k <- 1
while (lat > S.most.lat){
  # find new latitude
  tmp.pt <- data.frame(destPoint(E.coords[k, ], 180, dlines))
  lat <- tmp.pt$lat
  new.line <- SpatialLines(list(Lines(list(Line(data.frame(long = c(W.end, E.end),
                                                           lat = c(lat, lat)))),
                                      ID = "a")),
                           proj4string = CRS("+proj=longlat +datum=WGS84"))
  
  new.coords <- coordinates(gIntersection(SDBay.gis, new.line))
  new.coords.df <- data.frame(new.coords)
  
  for (c in seq(1, ncol(new.coords.df), by = 2)){
    k <- k + 1
    W.coords[k, ] <- new.coords.df[1, c:(c+1)]
    E.coords[k, ] <- new.coords.df[2, c:(c+1)]
    
  }

}

W.coords %>% transmute(long.W = long, lat.W = lat) -> W.coords
E.coords %>% transmute( long.E = long, lat.E = lat) -> E.coords

# find distance between beginning and end:
dist.line <- distHaversine(W.coords, E.coords)

# get rid of those that are < min.length m, then reorder lat/lon so that I can copy and paste
# in Google Earth to plot end points easily (they like lat, long pairs, not long, lat)

all.coords <- cbind(W.coords, E.coords, dist.line)
all.coords %>% filter(dist.line > min.length) %>%
  filter(lat.W > S.most.lat) %>%
  transmute(lat.W = lat.W, long.W = long.W,
            lat.E = lat.E, long.E = long.E,
            length = dist.line) -> survey.coords

survey.coords$ID <- seq(1, nrow(survey.coords))

survey.coords %>% select(ID, lat.W, long.W, lat.E, long.E, length) -> survey.coords
```




Once W and E ends are created, plot them on a map.

```{r}
water.gis <- spTransform(readOGR(dsn = "GISfiles",
                                 layer = "water",
                                 verbose = FALSE),
                         CRS("+proj=longlat +datum=WGS84"))

water.df <- broom::tidy(water.gis)


# make a plot
p.1 <- ggplot() +
 geom_polygon(data = water.df,
               aes(x = long, y = lat, group = group),
               fill = water.color,
               color = "black") +
  geom_polygon(data = SDBay.df,
               aes(x = long, y = lat, group = group),
               fill = water.color,
               color = "black") +
  geom_segment(data = survey.coords,
               aes(x = long.W, y = lat.W,
                   xend = long.E, yend = lat.E),
               size = 0.5,
               color = "darkgreen") +
  geom_text(data = survey.coords,
            aes(x = long.W,
                y = lat.W, label = ID),
            color = 'red',
            size = 4)+
  coord_map(ylim = c(32.57, 32.68),
            xlim = c(-117.2, -117.08)) +
  xlab('') + ylab('')

p.1

if (save.image) {
  ggsave(plot = p.1,
         dpi = 600,
         file = paste0('figures/UAS_tracklines_', 
                       dlines, 'm_', Sys.Date(), '.png'))

  write.table(signif(survey.coords, digits = 8), sep = ',',
            file = paste0('data/UAS_tracklines_', 
                          dlines, 'm_', Sys.Date(), '.csv'),
            append = F, quote = F, row.names = F, col.names = T)

}

```

